##Is There No Making Some People Happy?

###Introduction
There are categories of businesses in the [Yelp! dataset](http://www.yelp.com/dataset_challenge) that may very well transcend the quality of the businesses within them. On average, the businesses within these categories enjoy very high ratings from Yelp! reviewers. Nevertheless, there are those individuals who have assigned to buisinesses within one of these **pleasure** categories at least one review that is at least 1.5 stars lower than the category's average rating. These people will heretofore be known as the **the miserables**. The inverse is also true: there are those who have assigned at least one rating at least 1.5 stars higher than the average category rating for a business in a **non-pleasure** category. They are **the joyous**.

Are there categories that defy one's status as **miserable** or **joyous**?

###Methods

Are the **miserables** and the **joyous** all that different? One way to tell from the Yelp! dataset is by the feedback they receive from other Yelp! users. I chose to look at total feedback (a summation of all feedback compliments plus fan and friend counts), hot and funny compliments and the numer of reviews these users write. Drop lines denoting the mean for miserables, the joyous and the total population are blue, red and green, respectively.

Here's an example of how I produced a sampling of the mean sample by taking the mean of 2,000 samples from each population 100,000 times.

```{r, eval=F}
joyous.out <- replicate( 100000, mean( sample(joyous$funny_compliments, 2000) ) )
miserables.out <- replicate( 100000, mean( sample(miserables$funny_compliments, 2000) ) )
avgs.out <- replicate( 100000, mean( sample(avgs$funny_compliments, 2000) ) )
```

```{r,echo=F}
suppressMessages(library(ggplot2))
library(RColorBrewer)
suppressMessages(library(dplyr))
g <- function(dat,dat.avgs, t = 'yests', x.label='x label') {
  ggplot(dat, aes(x=as.numeric(cnt))) + 
    geom_histogram(data=filter(dat,type == 'joy'),aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="black", fill="red", alpha=.1) +
    geom_vline(data=filter(dat,type == 'joy'),aes(xintercept=mean(as.numeric(cnt))),   # Ignore NA values for mean
               color=brewer.pal(8,"Set3")[4], linetype="dashed", size=2)+
    geom_density(data=filter(dat,type == 'joy'),alpha=.4, fill=brewer.pal(11,"Spectral")[1]) +
    geom_histogram(data=df.avgs,aes(y=..density..),      # Histogram with density instead of count on y-axis
                   binwidth=.5,
                   colour="pink", fill="orange", alpha=.1) +
    geom_vline(data=df.avgs,aes(xintercept=mean(as.numeric(cnt))),   # Ignore NA values for mean
               color=brewer.pal(8,"Set3")[7], linetype="dashed", size=2)+
    geom_density(data=df.avgs,alpha=.4, fill=brewer.pal(11,"Spectral")[6]) +
    geom_histogram(data=filter(dat,type == 'misery'),aes(y=..density..),      # Histogram with density instead of count on y-axis
                    binwidth=.5,
                    colour="orange", fill="blue", alpha=.1) +
    geom_density(data=filter(dat,type == 'misery'),alpha=.4, fill=brewer.pal(12,"Paired")[2]) +
    geom_vline(data=filter(dat,type == 'misery'),aes(xintercept=mean(as.numeric(cnt))),   # Ignore NA values for mean
               color=brewer.pal(8,"Set3")[3], linetype="dashed", size=2) + theme_bw() + ggtitle(t) + xlab(x.label)
}

df.joyous.out <- read.table('../joyous.out.tsv',sep='\t',header=T,stringsAsFactors = F)
names(df.joyous.out)[1]<-'review.cnt'
df.miserables.out <- read.table('../miserables.out.tsv',sep='\t',header=T,stringsAsFactors = F)
names(df.miserables.out)[1]<-'review.cnt'
df.avgs.out <- read.table('../avgs.out.tsv',sep='\t',header=T,stringsAsFactors = F)
names(df.avgs.out)[1]<-'review.cnt'

joyous.total <- read.table('../joyous.tsv',sep='\t',stringsAsFactors = F,header=T)
miserables.total <- read.table('../miserables.tsv',sep='\t',stringsAsFactors = F,header=T)
avgs.total <- read.table('../avgs.tsv',sep='\t',stringsAsFactors = F,header=T)
```

```{r,echo=F, fig.height=3}
joyous.out <- df.joyous.out$review.cnt
miserables.out <- df.miserables.out$review.cnt
avgs.out <- df.avgs.out$review.cnt
# summary(joyous.out)
# summary(miserables.out)
dat <- NA
dat <- cbind('joy', joyous.out)
names(dat) <- c('type','cnt')
dat2 <- NA
dat2 <- cbind('misery', miserables.out)
names(dat2) <- c('type', 'cnt')
dat <- data.frame(rbind(dat,dat2), stringsAsFactors = F)
names(dat) <- c('type','cnt')
df.avgs <- data.frame(as.numeric(avgs.out), stringsAsFactors = F)
names(df.avgs) <- 'cnt'

plot1 <- g(dat,df.avgs, t='Review Counts',x.label='Means of Review Counts')

#feedback
joyous.out <- df.joyous.out$feedback
miserables.out <- df.miserables.out$feedback
avgs.out <- df.avgs.out$feedback
dat <- NA
dat <- cbind('joy', joyous.out)
names(dat) <- c('type','cnt')
dat2 <- NA
dat2 <- cbind('misery', miserables.out)
names(dat2) <- c('type', 'cnt')
dat <- data.frame(rbind(dat,dat2), stringsAsFactors = F)
names(dat) <- c('type','cnt')
df.avgs <- data.frame(as.numeric(avgs.out), stringsAsFactors = F)
names(df.avgs) <- 'cnt'

plot2 <- g(dat,df.avgs, t='Total Feedback',x.label='Means of Total Feedback')

#hot
joyous.out <- df.joyous.out$hot
miserables.out <- df.miserables.out$hot
avgs.out <- df.avgs.out$hot
dat <- NA
dat <- cbind('joy', joyous.out)
names(dat) <- c('type','cnt')
dat2 <- NA
dat2 <- cbind('misery', miserables.out)
names(dat2) <- c('type', 'cnt')
dat <- data.frame(rbind(dat,dat2), stringsAsFactors = F)
names(dat) <- c('type','cnt')
df.avgs <- data.frame(as.numeric(avgs.out), stringsAsFactors = F)
names(df.avgs) <- 'cnt'

plot3 <- g(dat,df.avgs,t='Hot Compliments',x.label='Means of Hot Compliments')

#funny
joyous.out <- df.joyous.out$funny
miserables.out <- df.miserables.out$funny
avgs.out <- df.avgs.out$funny
dat <- NA
dat <- cbind('joy', joyous.out)
names(dat) <- c('type','cnt')
dat2 <- NA
dat2 <- cbind('misery', miserables.out)
names(dat2) <- c('type', 'cnt')
dat <- data.frame(rbind(dat,dat2), stringsAsFactors = F)
names(dat) <- c('type','cnt')
df.avgs <- data.frame(as.numeric(avgs.out), stringsAsFactors = F)
names(df.avgs) <- 'cnt'

plot4 <- g(dat,df.avgs, t='Funny Compliments',x.label='Means of Funny Compliments')

suppressMessages(require(gridExtra))
grid.arrange(plot2, plot3, ncol=2,nrow=1)
```

```{r, echo=F,fig.height=3}
grid.arrange(plot4, plot1, ncol=2,nrow=1)
```

It's clear that these groups are perceived differently. And Student's t-Tests confirm these means are indeed different. Of interest though is that though the miserables receive less total feedback and fewer hot and funny compliments than the overall population, they write considerably more reviews than the overall population, though fewer than the joyous.

```{r, eval=F}
t.test(as.numeric(df.joyous.out$feedback), as.numeric(df.miserables.out$feedback))$p.value
t.test(as.numeric(df.joyous.out$hot), as.numeric(df.miserables.out$hot))$p.value
t.test(as.numeric(df.joyous.out$funny), as.numeric(df.miserables.out$funny))$p.value
```

```{r,echo=F}

suppressMessages(library(tm))
dest <- 'categories_nlp_docs'
setwd('/home/ubuntu/coursera_capstone/')

# create corpus
docs <- Corpus(DirSource(dest,pattern="tsv"));rm(dest)
# remove numbers
docs <- tm_map(docs, removeNumbers)

# remove stop words
docs <- tm_map(docs, removeWords, stopwords("SMART"))

# make all letters lower case
docs <- tm_map(docs, content_transformer(tolower))
# remove numbers
docs <- tm_map(docs, removeNumbers)

# remove stop words
docs <- tm_map(docs, removeWords, stopwords("SMART"))

# make all letters lower case
docs <- tm_map(docs, content_transformer(tolower))
library(SnowballC)
docs <- tm_map(docs, removePunctuation)
docs <- tm_map(docs, stripWhitespace)

docs <- tm_map(docs, stemDocument)

dtm <- DocumentTermMatrix(docs)
freq <- colSums(as.matrix(dtm))
ord <- order(freq)


# most frequent terms
#freq[tail(ord,n=25)]
dtms <- removeSparseTerms(dtm, 0.3)

# make a plot of freq terms with correlation above .6
# plot(dtms,terms =findFreqTerms(dtms, lowfreq=50))
```

```{r,echo=F,results='hide'}
df <- data.frame(inspect(TermDocumentMatrix(docs)))
```

Here are some examples of words more commonly found in **pleasurable** categories than **unpleasurable** categories. Not surprisingly, these words are somewhat affirming -- active, life, beauty.

```{r,echo=F, fig.height=1}
suppressMessages(library(dplyr))
df$diff <- df$joy.tsv - df$misery.tsv
df$pct.diff <- df$diff/(df$joy.tsv + df$misery.tsv)
df$names <- row.names(df)
df.pleasure <- head(arrange(filter(df, misery.tsv > 0), desc(diff)))
names(df.pleasure) <- c('pleas.category','unpleas.category','difference','pct.difference','name')
df.pleasure
```

It should be noted that to be considered among the **pleasure** categories, a category must include at least seven businesses and at least 30 reviews. The same is true for inclusion in the **unpleasurable** categories. And words more likely to appear in those can be found below:

```{r,echo=F}
df.unpleasure <- head(arrange(filter(df, misery.tsv > 0), diff))
names(df.unpleasure) <- c('pleas.category','unpleas.category','difference','pct.difference','name')
df.unpleasure

```

```{r, echo=F}
suppressMessages(library(caret))
df.dtm <- read.table("../dtm.csv",sep=',',header=T,stringsAsFactors = F)
cats <- read.table('../cats.tsv',sep='\t',header=T, stringsAsFactors = F)
df.dtm <- cbind(cats$categories, df.dtm[2:nrow(df.dtm),])
names(df.dtm)[1]<-"categories"
df.dtm$categories <- as.character(df.dtm$categories)
df.dtm <- df.dtm[,-2]
avg.stars <- read.table('../avg.stars.tsv',sep='\t',stringsAsFactors = F,header = T)
df.stars <- inner_join(avg.stars,df.dtm,by='categories')

rm(df.dtm);rm(avg.stars);rm(m);rm(cats)

set.seed(666)
trainIndex <- createDataPartition(df.stars$avg_stars, p = .8, list=F)
train <- df.stars[trainIndex,]
test <- df.stars[-trainIndex,]

fit <- lm(avg_stars ~ .,train[,c(2,3,5:ncol(train))])

coefs <- data.frame(summary(fit)$coef)
coefs$name <- row.names(coefs)
names(coefs) <- c('estimate','se','t.val','p.val', 'name')
prime.coefs <- arrange(filter(coefs, p.val<.05),p.val)

train <- train[,colnames(train)%in%c('categories', 'avg_stars', prime.coefs$name)]
test <- test[,colnames(test)%in%c('categories', 'avg_stars', prime.coefs$name)]

fit2 <- lm(avg_stars ~ ., train[,2:ncol(train)])

coefs <- data.frame(summary(fit2)$coef)
coefs$name <- row.names(coefs)
names(coefs) <- c('estimate','se','t.val','p.val', 'name')
prime.coefs <- arrange(filter(coefs, p.val<.05),p.val)

train <- train[,colnames(train)%in%c('categories', 'avg_stars', prime.coefs$name)]
test <- test[,colnames(test)%in%c('categories', 'avg_stars', prime.coefs$name)]

fit3 <- lm(avg_stars ~ ., train[,2:ncol(train)])

coefs <- data.frame(summary(fit3)$coef)
coefs$name <- row.names(coefs)
names(coefs) <- c('estimate','se','t.val','p.val', 'name')
prime.coefs <- arrange(filter(coefs, p.val<.05),p.val)

train <- train[,colnames(train)%in%c('categories', 'avg_stars', prime.coefs$name)]
test <- test[,colnames(test)%in%c('categories', 'avg_stars', prime.coefs$name)]

fit4 <- lm(avg_stars ~ ., train[,2:ncol(train)])


plot(df.stars$hair, df.stars$avg_stars, type = "n", frame = FALSE)
# fit.hair <- lm(avg_stars~hair,df.stars)
# abline(fit.hair, lwd = 2, col='pink')
# abline(h = mean(df.stars[df.stars$hair!=0,]$avg_stars), lwd = 3)
# abline(h = mean(df.stars[df.stars$hair==0,]$avg_stars), lwd = 3, col='yellow')
fit.hair.feedback_quadrant <- lm(avg_stars~ hair * feedback_quadrant, df.stars)
abline(coef(fit.hair.feedback_quadrant)[1], coef(fit.hair.feedback_quadrant)[2], lwd = 3, col='red')
abline(coef(fit.hair.feedback_quadrant)[1] + coef(fit.hair.feedback_quadrant)[3], coef(fit.hair.feedback_quadrant)[2] + coef(fit.hair.feedback_quadrant)[4], lwd = 3, col='purple')
points(df.stars[df.stars$hair!=0,]$hair, df.stars[df.stars$hair!=0,]$avg_stars, pch = 21, col = "black", bg = "lightblue", cex = 2)
points(df.stars[df.stars$hair==0,]$hair, jitter(df.stars[df.stars$hair==0,]$avg_stars, 14), pch = 21, col = "black", bg = "salmon", cex = 2)
summary(fit.hair.feedback_quadrant)

plot(df.stars$restaur, df.stars$avg_stars, type = "n", frame = FALSE)
# fit.restaur <- lm(avg_stars~restaur,df.stars)
# abline(fit.restaur, lwd = 2, col='pink')
# abline(h = mean(df.stars[df.stars$restaur!=0,]$avg_stars), lwd = 3)
# abline(h = mean(df.stars[df.stars$restaur==0,]$avg_stars), lwd = 3, col='yellow')
fit.restaur.feedback_quadrant <- lm(avg_stars~ restaur * feedback_quadrant, df.stars)
abline(coef(fit.restaur.feedback_quadrant)[1], coef(fit.restaur.feedback_quadrant)[2], lwd = 3, col='red')
abline(coef(fit.restaur.feedback_quadrant)[1] + coef(fit.restaur.feedback_quadrant)[3], coef(fit.restaur.feedback_quadrant)[2] + coef(fit.restaur.feedback_quadrant)[4], lwd = 3, col='purple')
points(df.stars[df.stars$restaur!=0,]$restaur, df.stars[df.stars$restaur!=0,]$avg_stars, pch = 21, col = "black", bg = "lightblue", cex = 2)
points(df.stars[df.stars$restaur==0,]$restaur, jitter(df.stars[df.stars$restaur==0,]$avg_stars, 14), pch = 21, col = "black", bg = "salmon", cex = 2)
summary(fit.restaur.feedback_quadrant)

```

###Results
Upon producing a normalized distribution of sample means, I can compare more clearly the mean values for people in each of the populations. The joyous are represented by the red vertical line, the miserables' mean is denoted by the blue line and the mean for the total population is marked with a green drop line.


Using a Student's t-Test, it can be demonstrated that between each population, the population means indeed differ. And in each, case overall feedback is more positive for the joyous population.


###Discussion
One potential confounding variable that occurred to me is that it's possible the joyous just write more reviews than the miserables, and therefore receive more positive feedback by virture of being more active users. And though it is true (see below), both the joyous and miserables write more reviews than does the average population.
